Stateless Model Checking with Data-Race Preemption Points

### 2015-10-21 pldi-16/di.tex
finish describing quicksand implementation

With a limited CPU budget, we must avoid running tests likely to time out.
When a MC instance reports an ETA too high,
we compare other yet-untested PP sets to find another job more likely to complete in time.
This algorithm is the heart of Iterative Deepening; we list it in Algorithm~\ref{alg:shouldworkblock}\footnote{
Though its worst-case performance is $O(mn)$ in the number of pending and suspended jobs,
in practice the non-constant portion runs very infrequently
and is negligible compared to the exponentially-sized state spaces.}.
Our method's main feature is understanding that if \GetPPSet{$j_1$} $\subset$ \GetPPSet{$j_2$},
and $j_1$ is suspended,
then $j_2$'s state space is guaranteed to be strictly larger, so $j_2$ will take at least as long.
Hence we should avoid testing $j_2$ unless further time spent on $j_1$ reveals that it might finish in time after all.
Similarly, when some job finds a bug, we cancel all pending superset jobs, as they would find the same bug.

@@

\quicksand~is an independent program that wraps the execution of several \landslide~instances.
The implementation is roughly 3000 lines of C.
%Figure~\ref{fig:quicksand} depicts the major components.
%and a registry of PPs to keep track of data race candidates.

\quicksand~uses a workqueue model
...
priorities... data race priorities...

@@

{\bf Heuristics.}

@@

When \quicksand~receives a data race report, it adds two new jobs to its workqueue:
a ``small'' job to preempt on the racing instruction only,
and a ``big'' job to preempt on that instruction as well as each PP used by the reporting job.
%
Hence each {\em pair} of racing accesses will spawn four new jobs, as shown in Figure~\ref{fig:new-dr-jobs}.
Rather than adding a maximal job with both new PPs at once, we prefer to add smaller jobs which have a higher chance of completing in time.
If those state spaces are bug-free, they will in turn add the maximal job later.
%
The rationale of spawning multiple jobs is that which will be more fruitful cannot be known in advance:
while the big job risks not completing in time,
the small job risks missing the data race entirely if the original PPs were required to expose it.
In practice, we observed some bugs found quickly by these small jobs, and other bugs missed by the small jobs found eventually by the big jobs,
which motivates the need for Iterative Deepening to decide at runtime which to prioritize.
% TODO: Put numbers here.
% which justifies this design

@@

% TODO: make new-dr-jobs diagram

%                                  (2)          landslide         landslide
%  landslide                   "Preempt when-->  o  PPs=[foo]      o  PPs=[mutex_lock,
%    o  PPs=[mutex_lock]        PC = FOO"       / \               / \      foo]
%   / \                              |         o   o             o   o
%  o  /\                             |                          /|\  |\
%    o  o                                                      o o o o o
%                             __QUICKSAND__
%    |       (1)             |             |           (3)
%    \  "Hey, an access      |  job1       | ----"Preempt when -->
%     '- at FOO raced w/ --> |        job2 |      PC = BAR"
%        one at BAR!         |    job4     |
%                            |        job70|      [ same pictures for s/foo/bar/]
%                            | job8        |
%                            '-------------'
%                maybe show some internal structore of QS as well

>>>>

With a limited CPU budget, we must avoid running tests likely to time out.
When a MC instance reports an ETA too high,
we compare other yet-untested PP sets to find another job more likely to complete in time.
This algorithm is the heart of Iterative Deepening; we list it in Algorithm~\ref{alg:shouldworkblock}\footnote{
Though its worst-case performance is $O(mn)$ in the number of pending and suspended jobs,
in practice the non-constant portion runs very infrequently
and is negligible compared to the exponentially-sized state spaces.}.
Our method's main feature is understanding that if \GetPPSet{$j_1$} $\subset$ \GetPPSet{$j_2$},
and $j_1$ is suspended,
then $j_2$'s state space is guaranteed to be strictly larger, so $j_2$ will take at least as long.
Hence we should avoid testing $j_2$ unless $j_1$'s ETA improves after further time spent on it.
%reveals that it might finish in time after all.
Similarly, when some job finds a bug, we cancel all pending superset jobs, as they would find the same bug.

@@

\quicksand~is an independent program that wraps the execution of several \landslide~instances.
The implementation is roughly 3000 lines of C.
It uses a thread pool to schedule the available state spaces,
sorting such jobs according to their status among a running queue, pending queue, and suspended queue.
Jobs are further prioritized by number of PPs, ETA, and whether they include data-race PPs.

@@

{\bf Heuristics.}
% List of all heuristix:
% HOMESTRETCH - last 60sec of test, don't suspend
% ETA_THRESH - "to let its ETA stabilize"
% eta factor
% shold_reproduce -- small dr jobs are not allowed to add further instances of themself (why not? don't remember)
% priority change between suspected and confirmed dr
Algorithm~\ref{alg:shouldworkblock} allows heuristically scaling a job's ETA when comparing to the time budget,
to express how optimistic we are about the estimate's accuracy.
We use 2 as this scaling factor based on the results in \cite{estimation},
though we allow changing it via the command line.
We also include a heuristic to
%ignore ETAs entirely
never suspend jobs before they pass a certain threshold of interleavings tested,
for which we choose 32,
so their ETAs have some time to stabilize.

We classify data-race candidates as {\em single-order} or {\em both-order} \cite{portend}
based on whether \landslide~observed the access pair ordered one or both ways while testing the original state space,
Single-order candidates are much more likely to be false-positives
(though we cannot say for sure before preempting during the access itself, hence the heuristic),
so we prioritize jobs with both-order data-race PPs more highly.
For single-order races, we do not add PPs for the later access at all;
should they be needed, preempting on the first access will suffice to upgrade it to both-order.

@@

When \quicksand~receives a data race report, it adds two new jobs to its workqueue:
a ``small'' job to preempt on the racing instruction only,
and a ``big'' job to preempt on that instruction as well as each PP used by the reporting job.
%
Hence each {\em pair} of racing accesses will spawn four new jobs, as shown in Figure~\ref{fig:new-dr-jobs}.
Rather than adding a maximal job with both new PPs at once, we prefer to add smaller jobs which have a higher chance of completing in time.
If those state spaces are bug-free, they will in turn add the maximal job later.
%
The rationale of spawning multiple jobs is that which will be more fruitful cannot be known in advance:
while the big job risks not completing in time,
the small job risks missing the data race entirely if the original PPs were required to expose it.
In practice, we observed some bugs found quickly by these small jobs, and other bugs missed by the small jobs found eventually by the big jobs,
which motivates the need for Iterative Deepening to decide at runtime which to prioritize.
% TODO: Put numbers here.

@@
